---
title: "CBB Conference Index"
format: html
server: shiny
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
# Load library because of trouble sorting out cbb_logo_links
library(cbbplotR)
```

```{r data-load, warning=FALSE, include=FALSE, message=FALSE, warning=FALSE}
#| context: setup
source(here::here("R/conf_diff_data.R"))

source(here::here("R/conf_net_summary.R"))

# Not including head-to-head charts right now
# source(here::here("R/conf_head_to_head.R"))
```

::: panel-tabset
### Standings
```{r conf-standings}
fluidPage(
  fluidRow(
    column(
      4,
      selectInput(
        'conf',
        'Select a conference',
        choices = c(as.character(unique(conf_margin$conf))),
        selected = "ACC"
      )
    )
  ),
  fluidRow(
    column(
      12,
      gt::gt_output("conftable")
    )
  )
)
```

### NET Summary

```{r net-summary}
fluidPage(
  fluidRow(
    column(
      4
      )
    ),
  fluidRow(
    column(
      10,
      gt::gt_output("by_conftable")
    )
  )
)
```


### About

Welcome to the college basketball conference index. This project was built by me, [Chris Gallo](https://hellogallo.com). All code is written in R.  


The goal of this project is to share tables of data about the 2023-24 men's college basketball season. This project is built on top of [cbbdata](https://github.com/andreweatherman/cbbdata) and [cbbplotR](https://cbbplotr.aweatherman.com/index.html) both developed by [Andrew Weatherman](https://www.aweatherman.com/). 

There are two types of tables surfacing the following data: 

- Conference standings and point differentials by location
- NET summary by conference 

**Standings**

The standings show win/loss records and point differentials by location (home or away). You can filter by any conference to see the specific data. 

The intention is to show the raw differential only, so you can quickly grasp the performance of certain teams in conference play only. 

If you're looking for adjusted metrics or standngs, you can check out other data sources like [barttorvik.com](https://barttorvik.com/#) or [kenpom.com](https://kenpom.com/) or [John Gasaway](https://johngasaway.com/).

**NET Summary**

The NET Summary table is a static table that surfaces the home win percentage in conference play, the average NET ranking of teams in a single conference, and the percentage of conference games in [each quadrant](https://www.ncaa.com/news/basketball-men/article/2022-12-05/college-basketballs-net-rankings-explained). 

This can be used as a gauge of quadrant opportunities in conference play. For example, ~60 percent of Big 12 games are in the first quadrant. 

**Data updates**

The goal is to update this data daily, please stay tuned. 

**Thank you**

If you found this useful, please consider sharing it. I also write a data visualization newsletter, [Bless your chart](https://blessyourchart.substack.com/), about college sports. 

:::

```{r shiny-server}
#| context: server
output$conftable <- gt::render_gt({
    cached_conf_deltas(conf = input$conf)
})


output$by_conftable <- gt::render_gt({
    by_conf_table
})
```
